---
title: "One Microservice - A simulation in simmer"
author: Patrick Boehnke
date: "December 30, 2024"
output:
  html_document:
    number_sections: true
---

# Introduction
We're going to develop a simulation of a single microservice using [simmer](https://r-simmer.org/). The purpose is two fold:

1. I want to learn more about using simmer
2. I want to build towards a model that incorporates Bayesian statistics and discrete event simulation to understand the behavior of microservice based systems

## simmer
I want to have a simulation where I can generate events (API requests) using arbitrary size and arrival time distributions while getting information about queue times and load. In looking for a software package that helps me with this simulation work, I had a few additional requirements:

1. I wanted something that integrates with R or Python because this whole project is developed in Posit Cloud
2. I need it to be somewhat performant because we will simulate large systems
3. Has to handle arbitrary size and arrival time distributions because API requests are anything but normal

Originally, I was going to use [SimPy](https://simpy.readthedocs.io/en/latest/) since it meets my requirements. However, at the time of this writing the Python support in Posit Cloud is in beta. So I kept looking until I found [simmer](https://r-simmer.org/), a discrete event simulation package for R. So let's give it a go and learn how to use it together.

## Ultimate modeling aim
The goal of this article is to start down a journey of simulating microservice performance and reliability by combining realistic workloads with first principles based simulations. We will start with one simple microservice (this article) and gradually build up a system of interacting services. We can then explore what controls performance and reliability as well as learning how to improve our design by incorporating resiliency patterns.

# Defining terms for our simulation
1. Event - this is a single API request to the service to perform a discrete unit of work
2. Service - an entity that takes Events and processes them consisting of a Compute layer and a Data Storage layer
3. Resource - a concept in simmer used to represent the individual layers of each Service

# Setup for simmer
## Installation
Installing simmer is quite easy and instructions can be found [here](https://r-simmer.org/#installation)

## Setup
To get started we will need to tell R to load simmer, set a random number generator seed, and start a simmer environment.

```{r}
library(simmer)

set.seed(42)

env <- simmer("OneMicroService")
env
```

# Writing our simulation

```{r}
request <- trajectory("request path") %>%
  ## add a api hosting app
  seize("app", 1) %>%
  timeout(function() rnorm(1, 10)) %>%
  release("app", 1) %>%
  ## add a backing database
  seize("database", 1) %>%
  timeout(function() rnorm(1, 5)) %>%
  release("database", 1)

env %>%
  add_resource("app", 1) %>%
  add_resource("database", 2) %>%
  add_generator("request", request, function() rnorm(1, 10, 2))

env %>% run(until = 400)

env %>% get_mon_arrivals() %>%
  transform(waiting_time = end_time - start_time - activity_time)
```

# Computing Environment
Presented for readers who want to run the presented code themselves.
```{r}
sessionInfo()
```